### 2021-02-26 rad-icfp-2021/simplediff.tex
fixing target unf

Now our program in UNF is in an equivalent form to straight-line programs. 
Before presenting our differentiation macro, we present the target language for this macro. 
We call this intermediate representation target UNF.
The main idea is similar to source UNF where we only have constants and sequential composition.
That said, the key difference is that now every constant represents a pair of an operation and what is meant to represent its Jacobian.
Following the insights from Section~\ref{sec:background}, the composition of pairs reverses the order on the second component. T
his represents the fact that reverse-mode reverts the computation flow, 
and that pre-composition (using a continuation) is the simplest way to do this in a purely functional setting.
On pairs, !;! still represents sequential composition for the first component, and reversed composition is represented by $\circ$ on the second component.
The grammar and types are given in Figure~\ref{fig:unf_target_grammar}.
For every constant $C$ from the source UNF, we have a new constant $J^TC$ representing its transpose Jacobian.

@@

The typesystem is given in Figure~\ref{fig:target_unf_typesystem}. 
One thing to note is that one can only form pairs.

>>>>

Now our program in UNF is in an equivalent form to straight-line programs. 
Before presenting our differentiation macro, we present the target language for this macro. 
We call this intermediate representation target UNF.
The main idea is similar to source UNF where we only have constants and sequential composition.
That said, the key difference is that now every constant represents a pair of an operation and what is meant to represent its Jacobian.
Following the insights from Section~\ref{sec:background}, the composition of pairs reverses the order on the second component. T
his represents the fact that reverse-mode reverts the computation flow, 
and that pre-composition (using a continuation) is the simplest way to do this in a purely functional setting.
On pairs, !;! still represents sequential composition for the first component, and reversed composition is represented by $\circ$ on the second component.
The grammar and types are given in Figure~\ref{fig:unf_target_grammar}. 
For every constant $C$ from the source UNF, we have a new constant $J^TC$ representing its transpose Jacobian.
One thing to note is that one can only form pairs in this language. 
Similarly to the source UNF, sequential composition is the only operation which is not a constant of the language.

@@

Technically, all our terms in target UNF should have an extra index $\rho$ which is a type from Source.
This will be the return type of the continuation for the jacobian accumulation. 
In other words, if we want the full gradient of a term $\Gamma\vdash$!e!: $\reals$, $\rho$ should be chosen to be $\Gamma$.
For every type !T!, let $\Dtype$(!T!)$:=$!T!$\times$(!T!->$\rho$). 
The typesystem is given in Figure~\ref{fig:target_unf_typesystem}.
